---
title: "National Park Visits"
output: ioslides_presentation
---

# Getting Started

## Load libraries
```{r}
library(tidyverse)
library(skimr)
library(paletteer)
```

## Get dataframe of palettes
```{r}
colors<-
  paletteer::palettes_d_names
```

## Set palette
```{r}
my_palette <-
  paletteer_d(package="ggsci", palette="category20_d3")
my_palette2 <-
  paletteer_d(package="ggsci", palette="category20_d3", 1)
```

## Read csv
```{r}
parks <- 
  readr::read_csv("National_Parks_Statistics.csv", skip=3, na=c("NA", "N/A", "#N/A"))
```

## Get variable names
```{r}
names(parks)
```

## Summarize data
```{r}
glimpse(parks)
```

## Map na's
```{r}
parks %>%
  purrr::map_df(~ sum(is.na(.))) %>%
  tidyr::gather(variables, num_nas) %>% 
  arrange(desc(num_nas))
```

## Rename variables
```{r}
parks_renamed <- parks %>%
  dplyr::rename(
            park_name = ParkName,
            unit_code = UnitCode,                     
            park_type = ParkType,
               region = Region,                         
                state = State,
                year = Year,                           
                month = Month,   
            rec_visits = RecreationVisits,
        non_rec_vists = NonRecreationVisits,
            rec_hours = RecreationHours,
         non_rec_hours = NonRecreationHours,
          conc_lodging = ConcessionerLodging,            
          conc_camping = ConcessionerCamping,
          tent_campers = TentCampers,
            rv_campers = RVCampers,
            backcountry = Backcountry,               
non_rec_overnight_stays = NonRecreationOvernightStays,
  misc_overnight_stays = MiscellaneousOvernightStays,
      park_name_total = ParkNameTotal,
      unit_code_total = UnitCodeTotal,    
      park_type_total = ParkTypeTotal,            
        regional_total = RegionTotal,    
          state_total = StateTotal,
            year_total = YearTotal,
      rec_visits_total = RecreationVisitsTotal,
  non_rec_visits_total = NonRecreationVisitsTotal,
      rec_hours_total = RecreationHoursTotal,
  non_rec_hours_total = NonRecreationHoursTotal,   
    conc_lodging_total = ConcessionerLodgingTotal,
    conc_camping_total = ConcessionerCampingTotal,
    tent_campers_total = TentCampersTotal,
      rv_campers_total = RVCampersTotal,      
      backgrounty_total = BackcountryTotal,                
      non_rec_overnight = NonRecreationOvernightStaysTotal,
misc_overnight_stays_total = MiscellaneousOvernightStaysTotal
          )
```

## Get rid of scientific notation
```{r}
options(scipen=999)
```

# Preliminary Analysis

## What is the most popular park type? 
- GGNRA in 1987 had the highest average number of visits 
- NRAs seem to be the most popular park type, at least recreationally
```{r echo=FALSE}
parks_renamed %>%
  group_by(park_type) %>%
   summarize(mean_rec_visits = mean(rec_visits), 
              min_rec_visits = min(rec_visits),
              max_rec_visits = max(rec_visits),
              total = n()) %>%
  arrange(desc(mean_rec_visits)) %>%
  ggplot(aes(x= reorder(park_type, mean_rec_visits), y=mean_rec_visits, fill=park_type)) +
  geom_bar(stat="identity", alpha = 0.7, color= "white") +
  labs(title = "Mean Rec Visits per Park Type",
       x = "Park Type",
       y = "Mean Rec Visits") +
  theme(legend.position = "none", axis.text.x = element_text(angle = 65, hjust = 1),
       plot.title=element_text(size=16, face="bold", hjust=0.5),
        axis.text = element_text(size=10),
        axis.title = element_text(size=12)) +
  scale_fill_manual(values=my_palette)
```

## What is the most popular park?
- Kobuk Valley is least popular
- Great Smoky Mountains is most popular
```{r echo=FALSE}
parks_renamed %>%
  filter(park_type=="National Park") %>%
  group_by(park_name) %>%
   summarize(mean_rec_visits = mean(rec_visits), 
              min_rec_visits = min(rec_visits),
              max_rec_visits = max(rec_visits),
              total = n()) %>%
  arrange(desc(mean_rec_visits)) %>%
  ggplot(aes(x=reorder(park_name, mean_rec_visits), y=mean_rec_visits)) +
  geom_bar(stat="identity", alpha = 0.7, fill=my_palette2) +
  labs(title = "Mean Rec Visits Per Park",
       x = "Park",
       y = "Mean Rec Visits") +
  theme(legend.position = "none", axis.text.x = element_text(angle = 65, hjust = 1),
       plot.title=element_text(size=16, face="bold", hjust=0.5),
       axis.text = element_text(size=6),
      axis.title = element_text(size=12))
```

```{r echo=FALSE}
graphaxis <- c(1,2,3,4,5,6,7,8,9,10,11,12)
```

## What is the best season to go?
- July is the most popular month
- January is least popular
```{r echo=FALSE}
parks_renamed %>%
  select(park_name, month, year, rec_visits) %>%
  filter(year > 2000) %>%
  group_by(month, year) %>%
  summarize(mean_rec_visits = mean(rec_visits)) %>%
  ggplot(aes(x=month, y=mean_rec_visits, group=month, fill=as.factor(month))) +
  geom_boxplot(alpha = 0.7) +
  labs(title = "Mean Rec Visits per Month",
       x = "Month",
       y = "Mean Rec Visits",
       fill="Month") +
  theme(axis.text.x = element_text(hjust = 1),
        plot.title=element_text(size=16, face="bold", hjust=0.5),
        axis.text = element_text(size=10),
        axis.title = element_text(size=12)) +
  scale_fill_manual(values=my_palette) +
  scale_x_continuous(breaks=graphaxis)
```

## What region is the most popular?
```{r}
parks_renamed %>%
  group_by(region) %>%
   summarize(mean_rec_visits = mean(rec_visits), 
              min_rec_visits = min(rec_visits),
              max_rec_visits = max(rec_visits),
              total = n()) %>%
  arrange(desc(mean_rec_visits))
```

## What region is the most popular?
```{r echo=FALSE}
parks_renamed %>%
  select(region, year, rec_visits) %>%
  filter(region != "National Capital") %>%
  group_by(year, region) %>%
  summarize(mean_rec_visits = mean(rec_visits)) %>%
  ggplot(aes(x=year, y=mean_rec_visits, group=region, color=region)) +
  geom_line() +
    labs(title = "Mean Rec Visits Per Region Over Time",
       x = "Year",
       y = "Mean Rec Visits") +
  theme(legend.position = "none", axis.text.x = element_text(hjust = 1),
       plot.title=element_text(size=16, face="bold", hjust=0.5),
        axis.text = element_text(size=6),
        axis.title = element_text(size=12)) +
  facet_wrap(~region) +
  scale_colour_manual(values=my_palette)
```

## How do regions compare?
```{r echo=FALSE}
parks_renamed %>%
  select(region, year, rec_visits) %>%
  group_by(year, region) %>%
  filter(region != "National Capital") %>%
  summarize(mean_rec_visits = mean(rec_visits)) %>%
  ggplot(aes(x=year, y=mean_rec_visits, group=region, color=region)) +
  geom_line() +
    labs(title = "Mean Rec Visits per Region Overtime",
       x = "Year",
       y = "Mean Rec Visits") +
  theme(axis.text.x = element_text(hjust = 1),
       plot.title=element_text(size=16, face="bold", hjust=0.5),
        axis.text = element_text(size=10),
        axis.title = element_text(size=12)) +
  scale_colour_manual(values=my_palette)
```

# Our question: How did the recessions of 1981 and 2008 affect rec visits in the 5 most popular National Parks?

## How did visits fare across the parks?
- visits have been going up going up over time
- peaked in late 80s then again in 2006 and onwards
- could suggest most popular parks become more popular, and least becoming less popular
```{r echo=FALSE}
parks_renamed %>%
  group_by(year) %>%
   summarize(mean_rec_visits = mean(rec_visits), 
              min_rec_visits = min(rec_visits),
              max_rec_visits = max(rec_visits),
              total = n()) %>%
  arrange(desc(mean_rec_visits)) %>%
  ggplot(aes(x=year, y=mean_rec_visits)) +
  geom_point() +
  geom_line() +
  labs(title = "Mean Rec Visits Overtime",
       x = "Year",
       y = "Mean Rec Visits") +
  theme(axis.text.x = element_text(hjust = 1),
       plot.title=element_text(size=16, face="bold", hjust=0.5),
        axis.text = element_text(size=10),
        axis.title = element_text(size=12)) +
  geom_vline(aes(xintercept=2008, colour="red")) +
  geom_vline(aes(xintercept=1981, colour="red")) +
  theme(legend.position = "none")
```

## How did visits fare in the Great Smoky Mountains?
```{r echo=FALSE}
parks_renamed %>%
  select(park_name, year, rec_visits) %>%
  filter(park_name=="Great Smoky Mountains NP") %>%
  group_by(year) %>%
  summarize(mean_rec_visits=mean(rec_visits))%>%
  ggplot(aes(x=year, y=mean_rec_visits)) +
  geom_point() +
  geom_line() +
  labs(title = "Mean Rec Visits to Great Smoky Mountains Overtime",
       x = "Year",
       y = "Mean Rec Visits") +
  theme(plot.title=element_text(size=16, face="bold", hjust=0.5),
        axis.text = element_text(size=10),
        axis.title = element_text(size=12)) +
  geom_vline(aes(xintercept=2008, colour="red")) +
  geom_vline(aes(xintercept=1981, colour="red")) +
  theme(legend.position = "none")
```

## How did visits fare in the Grand Canyon?
```{r echo=FALSE}
parks_renamed %>%
  select(park_name, year, rec_visits) %>%
  filter(park_name=="Grand Canyon NP") %>%
  group_by(year) %>%
  summarize(mean_rec_visits=mean(rec_visits))%>%
  ggplot(aes(x=year, y=mean_rec_visits)) +
  geom_point() +
  geom_line() +
  labs(title = "Mean Rec Visits to Grand Canyon NP Overtime",
       x = "Year",
       y = "Mean Rec Visits") +
  theme(plot.title=element_text(size=16, face="bold", hjust=0.5),
        axis.text = element_text(size=10),
        axis.title = element_text(size=12)) +
    geom_vline(aes(xintercept=2008, colour="red")) +
    geom_vline(aes(xintercept=1981, colour="red")) +
  theme(legend.position = "none")
```

## How did visits fare in Yosemite?
```{r echo=FALSE}
parks_renamed %>%
  select(park_name, year, rec_visits) %>%
  filter(park_name=="Yosemite NP") %>%
  group_by(year) %>%
  summarize(mean_rec_visits=mean(rec_visits))%>%
  ggplot(aes(x=year, y=mean_rec_visits)) +
  geom_point() +
  geom_line() +
  labs(title = "Mean Rec Visits to Yosemite NP Overtime",
       x = "Year",
       y = "Mean Rec Visits") +
  theme(plot.title=element_text(size=16, face="bold", hjust=0.5),
        axis.text = element_text(size=10),
        axis.title = element_text(size=12)) +
    geom_vline(aes(xintercept=2008, colour="red")) +
    geom_vline(aes(xintercept=1981, colour="red"))
  theme(legend.position = "none")
```

## How did visits fare in Yellowstone?
```{r echo=FALSE}
parks_renamed %>%
  select(park_name, year, rec_visits) %>%
  filter(park_name=="Yellowstone NP") %>%
  group_by(year) %>%
  summarize(mean_rec_visits=mean(rec_visits))%>%
  ggplot(aes(x=year, y=mean_rec_visits)) +
  geom_point() +
  geom_line() +
  labs(title = "Mean Rec Visits to Yellowstone NP Overtime",
       x = "Year",
       y = "Mean Rec Visits") +
  theme(plot.title=element_text(size=16, face="bold", hjust=0.5),
        axis.text = element_text(size=10),
        axis.title = element_text(size=12)) +
    geom_vline(aes(xintercept=2008, colour="red")) +
    geom_vline(aes(xintercept=1981, colour="red")) +
  theme(legend.position = "none")
```

## How did visits fare in Olympic?
```{r echo=FALSE}
parks_renamed %>%
  select(park_name, year, rec_visits) %>%
  filter(park_name=="Olympic NP") %>%
  group_by(year) %>%
  summarize(mean_rec_visits=mean(rec_visits))%>%
  ggplot(aes(x=year, y=mean_rec_visits)) +
  geom_point() +
  geom_line() +
  labs(title = "Mean Rec Visits to Olympic NP Overtime",
       x = "Year",
       y = "Mean Rec Visits") +
  theme(plot.title=element_text(size=16, face="bold", hjust=0.5),
        axis.text = element_text(size=10),
        axis.title = element_text(size=12)) +
    geom_vline(aes(xintercept=2008, colour="red")) +
    geom_vline(aes(xintercept=1981, colour="red")) +
  theme(legend.position = "none")
```

## How do the graphs compare?
```{r echo=FALSE}
parks_renamed %>%
  select(park_name, year, rec_visits) %>%
  group_by(year, park_name) %>%
  filter(park_name == "Great Smoky Mountains NP" | park_name == "Grand Canyon NP" | park_name == "Yosemite NP" | park_name == "Yellowstone NP" | park_name == "Olympic NP") %>%
  summarize(mean_rec_visits = mean(rec_visits)) %>%
  ggplot(aes(x=year, y=mean_rec_visits, group=park_name, color=park_name)) +
  geom_line() +
    labs(title = "Mean Rec Visits in Top 5 Parks Overtime",
       x = "Year",
       y = "Mean Rec Visits") +
  theme(axis.text.x = element_text(hjust = 1),
       plot.title=element_text(size=16, face="bold", hjust=0.5),
        axis.text = element_text(size=10),
        axis.title = element_text(size=12)) +
  geom_vline(aes(xintercept=2008, colour="red")) +
  geom_vline(aes(xintercept=1981, colour="red")) +
  scale_colour_manual(values=my_palette)
```

## How did visits fare in the least popular park, Kobuk Valley?
```{r echo=FALSE}
parks_renamed %>%
  select(park_name, year, rec_visits) %>%
  filter(park_name=="Kobuk Valley NP") %>%
  group_by(year) %>%
  summarize(mean_rec_visits=mean(rec_visits))%>%
  ggplot(aes(x=year, y=mean_rec_visits)) +
  geom_point() +
  geom_line() +
  labs(title = "Mean Rec Visits to Kobuk Valley NP Overtime",
       x = "Year",
       y = "Mean Rec Visits") +
  theme(plot.title=element_text(size=16, face="bold", hjust=0.5),
        axis.text = element_text(size=10),
        axis.title = element_text(size=12)) +
    geom_vline(aes(xintercept=2008, colour="red")) +
    geom_vline(aes(xintercept=1981, colour="red")) +
  theme(legend.position = "none")
```

## How did visits fare in Gates of the Arctic?
```{r echo=FALSE}
parks_renamed %>%
  select(park_name, year, rec_visits) %>%
  filter(park_name=="Gates of the Arctic NP & PRES") %>%
  group_by(year) %>%
  summarize(mean_rec_visits=mean(rec_visits))%>%
  ggplot(aes(x=year, y=mean_rec_visits)) +
  geom_point() +
  geom_line() +
  labs(title = "Mean Rec Visits to Gates of the Arctic Overtime",
       x = "Year",
       y = "Mean Rec Visits") +
  theme(plot.title=element_text(size=16, face="bold", hjust=0.5),
        axis.text = element_text(size=10),
        axis.title = element_text(size=12)) +
    geom_vline(aes(xintercept=2008, colour="red")) +
    geom_vline(aes(xintercept=1981, colour="red")) +
  theme(legend.position = "none")
```

## How did visits fare in Lake Clark?
```{r echo=FALSE}
parks_renamed %>%
  select(park_name, year, rec_visits) %>%
  filter(park_name=="Lake Clark NP & PRES") %>%
  group_by(year) %>%
  summarize(mean_rec_visits=mean(rec_visits))%>%
  ggplot(aes(x=year, y=mean_rec_visits)) +
  geom_point() +
  geom_line() +
  labs(title = "Mean Rec Visits to Lake Clark NP Overtime",
       x = "Year",
       y = "Mean Rec Visits") +
  theme(plot.title=element_text(size=16, face="bold", hjust=0.5),
        axis.text = element_text(size=10),
        axis.title = element_text(size=12)) +
    geom_vline(aes(xintercept=2008, colour="red")) +
    geom_vline(aes(xintercept=1981, colour="red")) +
  theme(legend.position = "none")
```

## How did visits fare in Isle Royale?
```{r echo=FALSE}
parks_renamed %>%
  select(park_name, year, rec_visits) %>%
  filter(park_name=="Isle Royale NP") %>%
  group_by(year) %>%
  summarize(mean_rec_visits=mean(rec_visits))%>%
  ggplot(aes(x=year, y=mean_rec_visits)) +
  geom_point() +
  geom_line() +
  labs(title = "Mean Rec Visits to Isle Royale NP Overtime",
       x = "Year",
       y = "Mean Rec Visits") +
  theme(plot.title=element_text(size=16, face="bold", hjust=0.5),
        axis.text = element_text(size=10),
        axis.title = element_text(size=12)) +
    geom_vline(aes(xintercept=2008, colour="red")) +
    geom_vline(aes(xintercept=1981, colour="red")) +
  theme(legend.position = "none")
```

## How did visits fare in Katmai?
```{r echo=FALSE}
parks_renamed %>%
  select(park_name, year, rec_visits) %>%
  filter(park_name=="Katmai NP & PRES") %>%
  group_by(year) %>%
  summarize(mean_rec_visits=mean(rec_visits))%>%
  ggplot(aes(x=year, y=mean_rec_visits)) +
  geom_point() +
  geom_line() +
  labs(title = "Mean Rec Visits to Katmai NP Overtime",
       x = "Year",
       y = "Mean Rec Visits") +
  theme(plot.title=element_text(size=16, face="bold", hjust=0.5),
        axis.text = element_text(size=10),
        axis.title = element_text(size=12)) +
    geom_vline(aes(xintercept=2008, colour="red")) +
    geom_vline(aes(xintercept=1981, colour="red")) +
  theme(legend.position = "none")
```

## How do the graphs compare?
```{r echo=FALSE}
parks_renamed %>%
  select(park_name, year, rec_visits) %>%
  group_by(year, park_name) %>%
  filter(park_name == "Kobuk Valley NP" | park_name == "Gates of the Arctic NP & PRES" | park_name == "Lake Clark NP & PRES" | park_name == "Isle Royale NP" | park_name == "Katmai NP & PRES") %>%
  summarize(mean_rec_visits = mean(rec_visits)) %>%
  ggplot(aes(x=year, y=mean_rec_visits, group=park_name, color=park_name)) +
  geom_line() +
    labs(title = "Mean Rec Visits in Bottom 5 Parks Overtime",
       x = "Year",
       y = "Mean Rec Visits") +
  theme(axis.text.x = element_text(hjust = 1),
       plot.title=element_text(size=16, face="bold", hjust=0.5),
        axis.text = element_text(size=10),
        axis.title = element_text(size=12)) +
  geom_vline(aes(xintercept=2008, colour="red")) +
  geom_vline(aes(xintercept=1981, colour="red")) +
  scale_colour_manual(values=my_palette)
```