---
title: "SophieDevan_Project_3"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(skimr)
```

```{r}
parks <- 
  readr::read_csv("/Users/sophieborison/Desktop/National_Parks_Statistics.csv", skip=3, na=c("NA", "N/A", "#N/A"))
```

```{r}
names(parks)
```

```{r}
glimpse(parks)
```

```{r}
parks %>%
  purrr::map_df(~ sum(is.na(.))) %>%
  tidyr::gather(variables, num_nas) %>% 
  arrange(desc(num_nas))
```

```{r}
parks_renamed <- parks %>%
  dplyr::rename(
            park_name = ParkName,
            unit_code = UnitCode,                     
            park_type = ParkType,
               region = Region,                         
                state = State,
                year = Year,                           
                month = Month,   
            rec_visits = RecreationVisits,
        non_rec_vists = NonRecreationVisits,
            rec_hours = RecreationHours,
         non_rec_hours = NonRecreationHours,
          conc_lodging = ConcessionerLodging,            
          conc_camping = ConcessionerCamping,
          tent_campers = TentCampers,
            rv_campers = RVCampers,
            backcountry = Backcountry,               
non_rec_overnight_stays = NonRecreationOvernightStays,
  misc_overnight_stays = MiscellaneousOvernightStays,
      park_name_total = ParkNameTotal,
      unit_code_total = UnitCodeTotal,    
      park_type_total = ParkTypeTotal,            
        regional_total = RegionTotal,    
          state_total = StateTotal,
            year_total = YearTotal,
      rec_visits_total = RecreationVisitsTotal,
  non_rec_visits_total = NonRecreationVisitsTotal,
      rec_hours_total = RecreationHoursTotal,
  non_rec_hours_total = NonRecreationHoursTotal,   
    conc_lodging_total = ConcessionerLodgingTotal,
    conc_camping_total = ConcessionerCampingTotal,
    tent_campers_total = TentCampersTotal,
      rv_campers_total = RVCampersTotal,      
      backgrounty_total = BackcountryTotal,                
      non_rec_overnight = NonRecreationOvernightStaysTotal,
misc_overnight_stays_total = MiscellaneousOvernightStaysTotal
          )
```

# want to summarize number of rec visits per park per year

### I dont want total number of entries, but total of rows for that park
```{r}
parks_renamed %>%
  group_by(park_name) %>%
  summarize(total = n())
```

### GGNRA in 1987 had the highest rec_visits 
### NRAs seem to be the most popular park type, at least recreationally
```{r}
parks_renamed %>%
  group_by(year, park_name) %>%
   summarize(mean_rec_visits = mean(rec_visits), 
              min_rec_visits = min(rec_visits),
              max_rec_visits = max(rec_visits),
              total = n()) %>%
  arrange(desc(mean_rec_visits)) %>%
  ggplot(aes(x=reorder(park_name, mean_rec_visits), y=mean_rec_visits, group=year, fill=year)) +
  geom_bar(stat="identity") + #width = 0.8, position = position_dodge(width = 0.9))
  labs(title = "Rec Visits Per Park Per Year",
       x = "Park",
       y = "Rec Visits") +
  theme(legend.position = "none", axis.text.x = element_text(angle = 65, hjust = 1),
       plot.title=element_text(size=16, face="bold", hjust=0.5),
       axis.text = element_text(size=2),
        axis.title = element_text(size=12))
```

### rec visits per year
### rec visits have been slowly going up going up over time
### peaked in late 80s then again in 2010 and onwards
```{r}
parks_renamed %>%
  group_by(year) %>%
   summarize(mean_rec_visits = mean(rec_visits), 
              min_rec_visits = min(rec_visits),
              max_rec_visits = max(rec_visits),
              total = n()) %>%
  arrange(desc(mean_rec_visits)) %>%
  ggplot(aes(x=year, y=mean_rec_visits)) +
  geom_point() +
  geom_line() +
  labs(title = "Rec Visits Per Year",
       x = "Year",
       y = "Mean Rec Visits") +
  theme(axis.text.x = element_text(angle = 65, hjust = 1),
       plot.title=element_text(size=16, face="bold", hjust=0.5),
        axis.text = element_text(size=10),
        axis.title = element_text(size=12))
```

# want to look at what park type is most popular

```{r}
parks_renamed %>%
  group_by(park_type) %>%
   summarize(mean_rec_visits = mean(rec_visits), 
              min_rec_visits = min(rec_visits),
              max_rec_visits = max(rec_visits),
              total = n()) %>%
  arrange(desc(mean_rec_visits)) %>%
  ggplot(aes(x= reorder(park_type,mean_rec_visits), y=mean_rec_visits)) +
  geom_bar(stat="identity") +
  labs(title = "Popularity of Park Type",
       x = "Park Type",
       y = "Mean Rec Visits") +
  theme(axis.text.x = element_text(angle = 65, hjust = 1),
       plot.title=element_text(size=16, face="bold", hjust=0.5),
        axis.text = element_text(size=10),
        axis.title = element_text(size=12))
```

# want to look at popularity of biomes overtime

```{r}
parks_renamed %>%
  group_by(region) %>%
   summarize(mean_rec_visits = mean(rec_visits), 
              min_rec_visits = min(rec_visits),
              max_rec_visits = max(rec_visits),
              total = n()) %>%
  arrange(desc(mean_rec_visits))
```

```{r}
p1 <- parks_renamed %>%
  select(region, year, rec_visits) %>%
  filter(region=="Northeast") %>%
  group_by(year) %>%
  summarize(mean_rec_visits = mean(rec_visits)) %>%
  ggplot(aes(x=year, y=mean_rec_visits)) +
  geom_point() +
  geom_line()
```

```{r}
p2 <- parks_renamed %>%
  select(region, year, rec_visits) %>%
  filter(region=="Pacific West") %>%
  group_by(year) %>%
  summarize(mean_rec_visits = mean(rec_visits)) %>%
  ggplot(aes(x=year, y=mean_rec_visits)) +
  geom_point() +
  geom_line()
```

```{r}
p3 <- parks_renamed %>%
  select(region, year, rec_visits) %>%
  filter(region=="Southeast") %>%
  group_by(year) %>%
  summarize(mean_rec_visits = mean(rec_visits)) %>%
  ggplot(aes(x=year, y=mean_rec_visits)) +
  geom_point() +
  geom_line()
```

```{r}
p4 <- parks_renamed %>%
  select(region, year, rec_visits) %>%
  filter(region=="Midwest") %>%
  group_by(year) %>%
  summarize(mean_rec_visits = mean(rec_visits)) %>%
  ggplot(aes(x=year, y=mean_rec_visits)) +
  geom_point() +
  geom_line()
```

```{r}
p5 <- parks_renamed %>%
  select(region, year, rec_visits) %>%
  filter(region=="Intermountain") %>%
  group_by(year) %>%
  summarize(mean_rec_visits = mean(rec_visits)) %>%
  ggplot(aes(x=year, y=mean_rec_visits)) +
  geom_point() +
  geom_line()
```

```{r}
p6 <- parks_renamed %>%
  select(region, year, rec_visits) %>%
  filter(region=="Alaska") %>%
  group_by(year) %>%
  summarize(mean_rec_visits = mean(rec_visits)) %>%
  ggplot(aes(x=year, y=mean_rec_visits)) +
  geom_point() +
  geom_line()
```

```{r}
p7 <- parks_renamed %>%
  select(region, year, rec_visits) %>%
  filter(region=="National Capital") %>%
  group_by(year) %>%
  summarize(mean_rec_visits = mean(rec_visits)) %>%
  ggplot(aes(x=year, y=mean_rec_visits)) +
  geom_point() +
  geom_line()
```

```{r}
p1 + p2 + p3 + p4 + p5 + p6 + p7
```

# want to look at best seasons to go

### I want the mean of the rec visits per month across the years

### July is the most popular month, January is least popular
```{r}
parks_renamed %>%
  select(park_name, month, year, rec_visits) %>%
  filter(year > 2000) %>%
  group_by(month, year) %>%
  summarize(mean_rec_visits = mean(rec_visits)) %>%
  ggplot(aes(x=month, y=mean_rec_visits, group=month, fill=month)) +
  geom_boxplot()
```

# I want to look at Joshua Tree in particular

```{r}
parks_renamed %>%
  select(park_name, year, rec_visits) %>%
  filter(park_name=="Joshua Tree NP") %>%
  group_by(year) %>%
  summarize(mean_rec_visits=mean(rec_visits))%>%
  ggplot(aes(x=year, y=mean_rec_visits)) +
  geom_point() +
  geom_line() +
  labs(title = "Visits to Joshua Tree NP Overtime",
       x = "Year",
       y = "Mean Rec Visits") +
  theme(plot.title=element_text(size=16, face="bold", hjust=0.5),
        axis.text = element_text(size=10),
        axis.title = element_text(size=12))
```


# What about Glacier National Park?

```{r}
parks_renamed %>%
  select(park_name, year, rec_visits) %>%
  filter(park_name=="Glacier NP") %>%
  group_by(year) %>%
  summarize(mean_rec_visits=mean(rec_visits))%>%
  ggplot(aes(x=year, y=mean_rec_visits)) +
  geom_point() +
  geom_line() +
  labs(title = "Visits to Glacier NP Overtime",
       x = "Year",
       y = "Mean Rec Visits") +
  theme(plot.title=element_text(size=16, face="bold", hjust=0.5),
        axis.text = element_text(size=10),
        axis.title = element_text(size=12))
```